---
title: "Unit 11 Takeaways"
author: "Sean Kennedy"
date: "2/29/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(orcutt)
library(tidyverse)

```

```{r 11.3.2-4}
data(airlog) # load from tswge package
airlog1 = artrans.wge(airlog,phi.tr=1)
airlog1.12 = artrans.wge(airlog1,phi.tr = c(rep(0,11),1))
ww = est.ar.wge(airlog1.12,p = 12)

plotts.sample.wge(ww$res)

ljung.wge(ww$res, K=24)
```

### TAKEAWAYS





### MODELING RESIDUALS

### TESTING RESIDUALS FOR WHITE NOISE

- Visual

- Ljung-Box test

  Null: autocorrelations are zero for all lags k
  Alt: at least one is non-zero

### TAKEAWAY OLS IS NOT SUFFICIENT :

- OLS Method for S/N is not good, fails to reject at too high of a rate. Terrible way to test - assumptions of OLS are not met by correlated time series data.


### TAKEAWAY COCHRANE ORCUTT (S/N Test):

- Searches for a linear signal (mean reversion)

- fit a linear regression to get estimates of a and b paramters. 

- transform by caculating the difference between the LM and the realization to estimate the correlated residuals

- fit AR1 to get phi_1

### TAKEAWAY SOMETHNG WRONG WITH C-O TEST :

- Fails to converge....

### TAKEAWAY DIFFERENCING TO TEST FOR UNIT ROOTS:

- By diffing data at k lags we can test for the presence of roots Bk 

- Afterwards, we can rinse repeat until we get white noise (hopefully)


