---
title: "Statistical Foundations for Data Science - HW9 - Sean Kennedy"
author: "Sean Kennedy"
date: "7/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sqldf)
library(tidyverse)
```

```{r, echo=FALSE}
critical_t = function(alpha, df, two_sided=TRUE){
  if(two_sided){
    t = qt(1-alpha/2, df = df)
  }
  else{
     t = qt(1-alpha, df = df)
  }
  return(t)  
}

p_val = function(t, df, sides=2){
  p = sides * pt(t, df, lower.tail = F)
  return(p)
}

```

## Load Data

```{r}

test_data = read.csv('TEST DATA.csv')
baseball_data = read.csv('Baseball_Data.csv')

```


## Question 2: Test Scores

  Assess relationship between math and science scores:
  
  {Math|Science}: Math = B0 + B1(Science)

```{r}
  linear_model = lm(data=test_data, formula = math ~ science)
  summary(linear_model)
  
```
According to the model - there will be in increase in math score of approximately 0.596 points for every increase in science score of 1 points.The estimated effect at the intercept is 21.7 - this is merely exrapolation due to no observations of science scores of zero existing in the set.

```{r}
plot(linear_model)

```

## Six Step for each parameter:
  
    - Intercept:
    
    H0: B0 = 0
    Ha: B0 <> 0
    
    Critical Values:
    
    N = 200
    df = 200 - k = 200 - 2 = 198
    alpha = 0.1 (95% CL)
    t_critical = 2.601 (2-sided)
  
    test stat: 7.879
    
    p-value: 2.15E-13
    
    Reject H0
    
    Strong statistical evidence that the intercept parameter B0 is not equal to zero. In layman's terms - there is no evidence to show that a zero on the science exam would lead to a zero on the math exam. This interpretation cannot be extended to a larger population since this was not a random sample. Causal inference is also not possible due to the nature of the observational study.
    
```{r}
df = 198
alpha = .01
t = critical_t(alpha, df)
p_b0 = p_val(7.879, df)
```

  - Slope parameter:
  
    H0: B1 = 0
    Ha: B1 <> 0
    
    *Same critical value analysis as above*
    
    test stastic: 11.437
    
    p-value: 1.39E-23
    
```{r}
p_b1 = p_val(11.437, df)
```

    Strong statistical evidence that the slope parameter B1 is not equal to zero. In layman's terms - there exists overwhelming evidence of a linear relationship between math and science scores. This interpretation cannot be extended to a larger population since this was not a random sample. Causal inference is also not possible due to the nature of the observational study.
  

  
  
## Confidence Intervals:

```{r}
b0_ste = 2.75429
b0_est = 21.70019
b0_cl = c(b0_est - t * b0_ste, b0_est + t * b0_ste) 

b1_ste =  0.05218
b1_est = 0.59681
b1_cl = c(b1_est - t * b1_ste, b1_est + t * b1_ste) 


confint(linear_model, level = 0.99)

```


## Bonus??? Why not


```{r}
baseball_lm = lm(data=baseball_data, formula = Wins ~ Payroll)
summary(baseball_lm)

```


## Plots

```{r}
plot(baseball_lm)

```
