---
title: "Homework 2 DDS - Sean Kennedy"
output: html_notebook
---

1. What is a basic workflow for reproducible research?
    + EDA -> hypothesis formulation --> hypothesis testing --> statistical analysis --> draw conclusions --> repeat :) 
2. What are five practical tips for making research reproducible?
    + document
    + test
    + collaborate
    + I would add more - but these three cover almost everything
3. Give an example of how you might implement each tip.
    + code comments
    + code assertions/unit tests
    + source control
4. Which one of these do you think will be the most difficult?
    + collaboration is always a challenge for me - I have worked on my own for too long
```{r}
str(airquality)

# Take only Temp columns
Temperature <- airquality$Temp
hist(Temperature)

Ozone <- airquality$Ozone

# Histogram with added parameters
hist(Temperature,
     main="Maximum daily temperature at La Guardia Airport",
     xlab="Temperature in degrees Fahrenheit",
     xlim=c(50,100),
     col="darkmagenta",
     freq=FALSE
)

# TODO: Assignment 2, Q2A: Complete the following code to yield a scatterplot with x as Month and y as Temp
# You're going to customize your plot slightly.  Use the help function to assist you if needed.
# Make the x label "Month" and the y label "Temperature"
# Finally, make the title of the plot "Temperature by Month" 
plot(x=airquality$Month, y=Temperature, xlab = 'Month', ylab = 'Temperature', main = 'Temperature By Month')


# TODO: Assignment 2, Q2B: Build a scatter plot with x as Temperature and y as Ozone 
# Complete the following code:
# Make the x label "Temperature" and the y label "Ozone",
# Make the title of the plot "Temperature vs Ozone"
plot(x=Temperature, y=Ozone, xlab='Temperature', ylab='Ozone', main='Ozone vs Temperature')
```
```{r pressure_data}
summary(pressure)
```

```{r pressure_plot, echo=FALSE}
plot(pressure, main='Pressure vs Temperature')
```
```{r pressure_plot2, echo=FALSE}
plot(pressure$pressure, pressure$temperature, main = 'Temperature vs Pressure', xlab='Pressure', ylab='Temperature')
```
