---
title: "Time Series Unit "
author: "Sean Kennedy"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
library(stats)
library(lubridate)
data('sunspot.classic')

```

```{r EXAMPLE_QUESTION}
factor.wge(phi = c(1.59, -.544, -.511, .222))
```

```{r 5.5.5}
## characteristic equation
## 1+0.1Z+0.82Z^2-.16Z^3=1+.2z

factor.wge(phi=c(-.1, -.8))
factor.wge(phi=c(-.2))



```

```{r 5.5.6}
## characteristic equation
## 1-1.95Z+1.9Z^2 = 1-1.95z+1.9z^2

factor.wge(phi=c(1.95, -1.9))
factor.wge(phi=c(1.95, -1.9))

```

```{r 5.5.10}
## characteristic equation
## 1-.1Z+.5Z^2-.7z^3 = 1-.72z+.8z^2

plotts.true.wge(phi = c(.1,-.5, .7), theta = c(.72, -.8))


```

```{r 5.7.3}
## characteristic equation
## 1-1.95Z+1.9Z^2 = 0
psi.weights.wge(c(1.95, -1.9), lag.max = 4)

```

```{r FOR_LIVE_SESSION}
#run AIC5 on sunspot data
plotts.sample.wge(sunspot.classic)
aic5.wge(sunspot.classic)

```

```{r WALMART}
walmart_data = read.csv('Walmart.csv')
item8 = walmart_data %>% filter(item == 50 & store==9)
sample = plotts.sample.wge(item8$sales)
aic5.wge(item8$sales)
#best model is an ARMA(4, 2), ARMA(5, 1) is a close second
#all have components of AR/MA models


```

```{r CALC_RHO_1}
## characteristic equation
## 1-0.8Z+0.5Z^2 = 0
theta_1 = 0.5
theta_2 = 0
rho1 =  (-theta_1 + theta_1*theta_2)/(1+theta_1^2+theta_2^2)
rho1
rho2 =  (-theta_2)/(1+theta_1^2+theta_2^2)
rho2
#psi.weights.wge(theta = c(theta_1, theta_2), lag.max=3)

#GLM: Xt = at - theta_1*a_t_minus_1 - theta_2*a_t_minus_2
plotts.true.wge(theta=c(theta_1, theta_2))
```

```{r GERNATE_AN_ARMA}
sample = plotts.true.wge(phi=c(0.2, -.1), theta=c(theta_1, theta_2))
#p = 2
#q = 2
#ARMA(2,2)

```
```{r}
flights = read.csv('SWADelay.csv')
aic5.wge(flights$arr_cancelled)

#models 1 and 3 are pure AR models (AR(1) and AR(2))
#models 2 and 5 are pure MA models (MA(2) and MA(1))
#model 4 is an ARMA(1, 1)

```




### TAKEAWAY 1 MA(q)

- All auto_correlations for k > q = 0
- Inverse spectral density peaks
- ALL MA MODELS ARE STATIONARY (make useful components but very rarely useful on their own)

### TAKEAWAY 2 MODEL EQUIVALENCY

- ARMA(0, q) = MA(q)
- ARMA(p, 0) = AR(p)

### TAKEAWAY 3 MA INVERTABILITY

- Must be invertable (roots must me outside unit circle)
- 2 models with the same autocorrelation structure (model multiplicity): one will be invertable, one will not be
(how does multiplicity relate to entropy?)


### TAKEAWAY 4 AR MODELS AS INFINITE SUM 

- All AR models can be represented as GLP (General Linear Process)

- Any AR model can be represented as an infinite sum of random noise terms much like the components of a wavefunction in QM
- 