---
title: "Unit 7 Takeaways"
author: "Sean Kennedy"
date: "2/16/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

```{r 7.6.7}

phi= c(0.4, -0.6, 0.8)
mu = 9
var = 3.49
psi = psi.weights.wge(phi=phi, lag.max=3)
sigma = sqrt(var)

interval_half = sigma*(sum(c(1,psi[1:length(psi)-1]**2)))**0.5

```

```{r 7.7.3}

delays = read.csv('SWADelay.csv')
delays = log(delays$arr_delay)
ar12_phi = c(0.44, 0.02, -0.12, 0.08, 0.0, 0.02, 0.06, -0.09, 0.06, 0.07, 0.02, 0.37)


arima_12_phi = c(0.34, 0.07, -0.11, 0.07, 0.01, 0.02, 0.06, -0.09, 0.05, 0.07, 0.02, 0.39)
arima_12_theta = c(-.12)

ar12_forecasts = fore.arma.wge(delays, phi=ar12_phi, lastn = T, n.ahead = 30)
arima_12_forecasts = fore.arma.wge(delays, phi=arima_12_phi, theta = arima_12_theta, lastn = T, n.ahead = 30)

actual = delays[(length(delays)-29):length(delays)]
err_ar12 = mean((ar12_forecasts$f-actual)**2)
err_arima_12 = mean((arima_12_forecasts$f-actual)**2)

mean(delays)
```

```{r 7.9.6}
woodward_phi = c(-.36, -.05, -.14, .04, .09, -.02, .02, .17, .03, -.1, -.38)
woodward_forecasts = fore.aruma.wge(delays, phi=woodward_phi, d=1, s=12, lastn = T, n.ahead = 36)
actual = delays[(length(delays)-35):length(delays)]
err_woodward = mean((woodward_forecasts$f-actual)**2)
```


```{r 7.10.2}
amazon = read.csv('AMZN.csv')
amazon = amazon$Adj.Close
forecasts = fore.sigplusnoise.wge(amazon, linear = TRUE, freq=0, max.p=5, n.ahead = 30, lastn=FALSE, limits=TRUE)
max(forecasts$f)
forecasts$ul[30]
forecasts$ll[30]
```



```{r FOR_LIVE_SESSION_1}
data('ss08')
aic5.wge(ss08, p=0:10, q=0:10, type='aic')


plotts.parzen.wge(ss08)
plotts.sample.wge(ss08)


```

```{r FOR_LIVE_SESSION_2}
phi = c(.7)
sigma = 3.73
mu = 15.6

psi = psi.weights.wge(phi=phi, lag.max = 2)
interval_half = sigma*(sum(c(1,psi[1:length(psi)-1]**2)))**0.5
```

```{r}
psi[1:length(psi)-1]
    
```

```{r FOR_LIVE_SESSION_3}
amtrack = read.csv('AmtrakPassengersMonthly.csv')$Ridership

phi1 = c(0.5511, 0.1680, -0.0145, 0.0651, 0.1388, -0.2966, 0.1539, 0.1270, -0.1815, 0.0364, 0.1456, 0.6287, -0.3832, -0.0199, -0.1679) 

phi2 =  c(-0.02709541,  0.74213105)
theta2 = c(-0.5844596,  0.3836931)

phi3 = c(0.306943)
theta3 = c(0.7431719)

forecast1 = fore.arma.wge(amtrack, phi=phi1, lastn=TRUE, n.ahead=12, limits=TRUE, plot=TRUE)$f

forecast2 = fore.arma.wge(amtrack, phi=phi2, theta=theta2, lastn=TRUE, n.ahead=12, limits=TRUE, plot=TRUE)$f


forecast3 = fore.arma.wge(amtrack, phi=phi3, theta=theta3, lastn=TRUE, n.ahead=12, limits=TRUE, plot=TRUE)$f

actual = amtrack[(length(amtrack)-11):length(amtrack)]
err_1 = mean((forecast1-actual)**2)
err_2 = mean((forecast2-actual)**2)
err_3 = mean((forecast3-actual)**2)


```

### model 1 has smallest ASE (i.e is best fit)



### TAKEAWAY 1

- ARIMA Forecasting
  
  Use fore.aruma.wge() for forecasting.
  Limits become unbounded as l increases
  A factor of (1 − B) does not forecast a trend. An order of
  d > 1 is required to forecast a trend.
  
### TAKEAWAY 2

ARIMA with Seasonality Forecasting
The forecast for step l is same as the last s value. Use
fore.aruma.wge() for forecasting.
Limits become unbounded as l increases

### TAKEAWAY 3

 A factor of (1 − B) does not forecast a trend. An order of
d > 1 is required to forecast a trend.
 (1 − B) (1 − Bs
) = at is called an airline model