---
title: "Unit 6 Takeaways"
author: "Sean Kennedy"
date: "2/11/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tswge)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

```{r 6.4.3}
#true autocorrelation at lag 5
#(1-.6B+.8B^2)(1-B)^2*X(t)=(1+.3B)*a(t)
x = gen.arima.wge(500, phi=c(0.6, -0.8), theta=c(-0.3), d=2, sn=37)
plotts.sample.wge(x)
ar_factor = factor.wge(phi=c(0.6, -0.8))
```

```{r 6.4.4-6.4.6}
#(1-.6B+.8B^2)(1-B)^2*X(t)=(1+.3B)*a(t)
x = gen.arima.wge(500, phi=c(0.6, -0.8), theta=c(-0.3), d=2, sn=35)
x_12 = artrans.wge(x, 1)
parzen.wge(x_12)
aic5.wge(x_12)

# I think the answer for 6.4.6 is wrong..... should be NO
```

### Takeaway 1: (1-B)^k Factor

  - Dominates the realization depending on value of k (higher k = more dominant)
  - should be removed with appropriate differencing (diff = k)
  - roots lie on the unit cirlce (z = 1)
  - mix of B^k factors


```{r 6.6.2-6.6.3}
#(1-.6B+.8B^2)(1-B)^2*X(t)=(1-.3B+.7B^2)*a(t)
x = gen.aruma.wge(500, phi=c(0.6, -0.8), theta=c(0.3, -.7), s=12, sn=37)
x_diff_12 = artrans.wge(x, c(rep(0, 11), 1))
parzen.wge(x_diff_12)
aic5.wge(x_diff_12)

```

### Takeaway 2: Factoring can show the underlying structure of the realization.

  - factoring the AR components can show if there are frequencies that coincide with known frequencies of seasonal    models (quarterly, monthly, weekly etc)


### Takeaway 3: (1-B^k) (seasonality) Factor

  - Dominates the realization, should be removed with appropriate differencing (diff = k)
  - roots lie on the unit cirlce (z = 1)
  - only contains B^k factor


```{r 6.6.6-6.6.7}
#(1+.5B-.2B^2+B^4-.5B^5+.2B^6)*X(t)=(1+.92B)*a(t)
factor.wge(phi=c(-.5, .2, 0, 1, .5, -.2))
#quarterly seasonal model


#(1+.3B-1.2B^2-.4b^3-.5B^5+B^12-.3B^13+1.2B^14+.4B^15)*X(t)=a(t)
factor.wge(phi=c(-.3, 1.2, .4, 0, .5, 0, 0, 0, 0, 0, 0, 1, .3, -1.2, -.4))
#monthly seasonal model
```

### Takeaway 4: ARUMA/ARIMA

  - ARUMA is a seasonal model (an ARIMA with d=0 and n=s) defined by (p, s, q)
  - ARIMA is combination of an AR, MA and S/N model defined by (p, d, q)

### For Live Session - Part 1

```{r LIVE_SESSION_1}
data('ss08')
plotts.sample.wge(ss08)

#sunspot data is clearly cyclical from the ACF plot - potentially an AR(2) with complex roots  
#lets check the AIC

aic5.wge(ss08)

#AIC suggests an ARMA(5, 1) as best fit (could be an AR5 that reduces to AR(2) and AR(1) components)

ss08_10 = artrans.wge(ss08, phi.tr = c(rep(0, 9), 1))
plotts.sample.wge(ss08_10)
```


### For Live Session Part 2

```{r FOR_LIVE_SESSION_PART_2}
#factor table for seasonal 7 model
factor.wge(phi=c(rep(0,6), 1))

```
```{r FOR_LIVE_SESSION_PART_3}
#model a
#(1-B4) Xt = (1+.3B) (1-.6B+.8B2)at
# Seasonal with d=0 (p=0, q=3)
phi_a = c(rep(0, 3), 1)
theta_a = c(.3, -.62, -.24)
a = gen.aruma.wge(500, s=4, theta=theta_a, sn=37)
a_4 = artrans.wge(a, phi=phi_a)
aic5.wge(a_4)
plotts.sample.wge(a_4)
#model b
#(1+.3B) Xt = (1-.6B+.8B2) at
# ARMA with p=1, q=2
phi_b = c(-.3)
theta_b = c(.6, -.8)
b = gen.arma.wge(500,theta=theta_b, phi=phi_b,  sn=37)
aic5.wge(b)
plotts.sample.wge(b)
#model b
#(1-.1B-.99B2 +.013B3 +.2078B4 +.0888B5 +.00864B6)(1-B)2(1-B12) Xt = (1-.6B+.8B2)at
# ARUMA d=1, s=12 p=6, q=2
phi_c = c(.1, .99, -.013, -.2078, -0.888, -.00864)
factor.wge(phi = phi_c)
theta_c = c(.6, -.8)
#c = gen.aruma.wge(500,theta=theta_c, phi=phi_c, s=12, d=1,  sn=37)
#aic5.wge(c)
## Can't get this one to work :(

```

```{r STOCK_DATA}
pzza = read.csv('pzza.csv', header=FALSE)
pzza_1 = artrans.wge(pzza$V1, phi.tr = c(1))
#diffed data has all the characteristics of white noise - autocorrelations > k=0 are zero
#an appropriate model could be signal-noise model with sigma_noise = realized variance of stock and a wandering 
#component of d = 1

aic5.wge(pzza_1)

#AIC confirms white noise and also suggests an AR(2) or an MA(2) model as best fit

```