---
title: "Sean_Kennedy_Homework_5_Stats"
author: "Sean Kennedy"
date: "June 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggpubr)
library(multcomp)
library(car)
library(sqldf)
```

## R Question 1 Problem 5.25 from Text (ANOVA)

```{r}
file_name = 'Unit 6 Handicap Data.csv'
raw_data = read.csv(file_name)
res.aov = aov(Score ~ Handicap, data=raw_data)
summary(res.aov)
```

## Let's plot the fuck out of this!

```{r education_data, echo=FALSE}
# tak a random sample for good measure
set.seed(1234)
dplyr::sample_n(education_data, 10)
education_data$LogIncome2005 = log(education_data$Income2005) 
levels(education_data$Educ)
#show some group stats
group_by(education_data, Educ) %>% summarise(
  count = n(),
  mean = mean(LogIncome2005, na.rm = TRUE),
  sd = sd(LogIncome2005, na.rm = TRUE)
)
ggboxplot(education_data, x='Educ', y='LogIncome2005', color = 'Educ',
          palette=palette(rainbow(length(levels(education_data$Educ)))),
          ylab = 'LogIncome2005', xlab = 'Educ'
          )

ggline(education_data, x = 'Educ', y = 'LogIncome2005', 
       add = c('mean_se'), #(other values include: mean_sd, mean_ci, median_iqr,jitter ....)
       ylab='LogIncome2005', xlab='Educ')
```

## Tukey Multiple Pairwise Comparisons

```{r}
TukeyHSD(res.aov)
```

## Let's calc the same thing using multcomp package 
  
    #(why doensn't this work?)
```{r}
#summary(glht(res.aov, linfct = mcp(Educ = 'Tukey')))

```

## pariwise t-test, because "Why not?"
  
  # creates a matrix of t-tests between classes
```{r}
pairwise.t.test(education_data$LogIncome2005, education_data$Educ,
                p.adjust.method = 'BH')

```

## Check homogeneity of variance assumption
```{r}
plot(res.aov, 1)
```

## Levene's test for homogeneity of variance
```{r}
leveneTest(LogIncome2005 ~ Educ, data=education_data)
```

## Relaxed Homogeneity Tests
  
  - Welch one-way
  
  - paired t-test w/o equality of sd assumption)

```{r}
oneway.test(LogIncome2005 ~ Educ, data = education_data)
pairwise.t.test(education_data$LogIncome2005, education_data$Educ, p.adjust.method = 'BH', pool.sd = FALSE)
```

## Check Normality Assumption

```{r}
# Q-Q plot of residuals
plot(res.aov, 2)

# Extract the residuals 
aov_residuals = residuals(object = res.aov)

#Shapiro-Wilke on residuals
shapiro.test(x = aov_residuals)
```

## Non Parametric Tests (Kruskal-Wallis Ranksum Test)

```{r}
kruskal.test(LogIncome2005 ~ Educ, data=education_data)

```

## R Question 2 BYO ANOVA

# Filter Data
```{r}
education_data_filtered = sqldf("SELECT * FROM education_data
                                              WHERE Educ IN ('16', '>16')")

```

# verify quoted p-value

```{r}
t.test(LogIncome2005 ~ Educ, data= education_data_filtered, var.equal=TRUE)

```