---
title: "Statistical Foundations for Data Science - HW10 - Sean Kennedy"
author: "Sean Kennedy"
date: "7/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sqldf)
library(tidyverse)
library(investr)
library(Sleuth3)
```

```{r, echo=FALSE}

df = 19
alpha = .01

critical_t = function(alpha, df, two_sided=TRUE){
  if(two_sided){
    t = qt(1-alpha/2, df = df)
  }
  else{
     t = qt(1-alpha, df = df)
  }
  return(t)  
}

p_val = function(t, df, sides=2){
  p = sides * pt(t, df, lower.tail = F)
  return(p)
}

#assumes 99% CL

```

## Load Data

```{r}
bird_data = read.csv('Male Display Data Set.csv')
summary(bird_data)

```


## Question 2: Stone Mass vs Immune Response

  Assess relationship between stone mass carried and immune response (t-cell):
  
  {t-cell|mass}: t-cell = B0 + B1(mass)

```{r}
  linear_model = lm(data=bird_data, formula = Tcell ~ Mass)
  summary(linear_model)
```
## Plot CL

```{r}
bird_plot = ggplot(bird_data, aes(x=Mass, y=Tcell)) + 
  geom_point(color='Blue', size = 2) + 
  geom_smooth(method = lm, se=TRUE, level=0.99, color='Red')

predictions_cl = predict(linear_model, newdata = data.frame(Mass = sort(bird_data$Mass)), interval = c('predict'), type=c('response'), level=0.99)

regress_cl = predict(linear_model, newdata = data.frame(Mass = sort(bird_data$Mass)), interval = c('confidence'), type=c('response'), level=0.99)

bird_plot + 
  geom_line(data=data.frame(predictions_cl), aes(x=sort(bird_data$Mass), y=upr), color='Green') +
  geom_line(data=data.frame(predictions_cl), aes(x=sort(bird_data$Mass), y=lwr), color='Green') +
  geom_line(data=data.frame(regress_cl), aes(x=sort(bird_data$Mass), y=upr), color='Blue') +
  geom_line(data=data.frame(regress_cl), aes(x=sort(bird_data$Mass), y=lwr), color='Blue')
```

## Six Step for each parameter:
  
    - Intercept:
    
    H0: B0 = 0
    Ha: B0 <> 0
    
    Critical Values:
    
    N = 21
    df = 21 - k = 21 - 2 = 19
    alpha = 0.01 (99% CL)
    t_critical = 2.861 (2-sided)
  
    test stat: 1.112
    
    p-value: 0.28
    
    Fail to reject H0
    
    There is not sufficient evidence to show that the intercept parameter B0 is not equal to zero (p-value = 0.28 at alpha of 0.01). Confidence intervals for this estimate are below. This interpretation cannot be extended to a larger population since this was not a random sample. Causal inference is also not possible due to the nature of the observational study.
    
```{r}

t = critical_t(alpha, df)
p_b0 = p_val(1.112, df)
```

  - Slope parameter:
  
    H0: B1 = 0
    Ha: B1 <> 0
    
    *Same critical value analysis as above*
    
    test stastic: 3.084
    
    p-value: 0.0061
    
```{r}
p_b1 = p_val(3.084, df)
```

    Strong statistical evidence that the slope parameter B1 is not equal to zero. In layman's terms - there exists strong evidence of a linear relationship between the ability to carry mass and health of the immune system. This interpretation cannot be extended to a larger population since this was not a random sample. Causal inference is also not possible due to the nature of the observational study. There may also be confounding variables that contribute to the ability to carry objects - namely the size of the brird. 
  


## Confidence Intervals:

```{r}
confint(linear_model, level = 0.99)

```

- Regression equation:

  Tmean = 0.0875 + 0.0328(Mass)

- Slope:

  For every 1g increase in mass the expected change in T-cell count is 0.0328
  
- Intercept:

  Baseline T-cell response = 0.0875 (the reponse when mass = 0, which sounds like a dead animal)
  

```{r}

prediction_at_target = predict(linear_model, newdata = data.frame(Tcell = NA, Mass=4.5), interval = "confidence", level=0.99)
prediction_at_target

```

## Extend analysis


```{r}
test_input = 0:25
test_y = rep(0.3, 26)

prediction_confidence = predict(linear_model, newdata = data.frame(Mass=test_input), interval = c("confidence"), type = c("response"), level =0.99) 

prediction_predict = predict(linear_model, newdata = data.frame(Mass=test_input), interval = c("predict"), type = c("response"), level =0.99) 


ggplot() + 
  geom_line(data=data.frame(prediction_confidence), aes(x=sort(test_input), y=upr), color='Green', lty=3) +
  geom_line(data=data.frame(prediction_confidence), aes(x=sort(test_input), y=lwr), color='Green', lty=3) +
  geom_line(data=data.frame(prediction_predict), aes(x=sort(test_input), y=upr), color='Blue', lty=3) +
  geom_line(data=data.frame(prediction_predict), aes(x=sort(test_input), y=lwr), color='Blue', lty=3) + 
  geom_line(data=data.frame(prediction_predict), aes(x=sort(test_input), y=fit), color='Red') + 
  geom_point(data=bird_data, aes(x=Mass, y=Tcell), color='Black') + 
  geom_line(aes(x=test_input, y=test_y), color='Purple', lty=2) + 
  geom_vline(aes(xintercept=22)) +
  xlab('Mass') + 
  ylab('T-Cell') +
  ggtitle('T-Cell vs Mass')

```

 *At a single response of 0.3 the prediction interval is from zero to 22g from inspection*

## Mean response
```{r}
calibrate(linear_model, y0=0.3, interval = 'inversion', mean.response = T, level = 0.99)

```

## Single Response
```{r}
calibrate(linear_model, y0=0.3, interval = 'inversion', level = 0.99)

```

## Interpretation

  - Mean response:
    
    @99% confidence - we estimate that to observe a mean t-cell repsonse of 0.3, the mass should be between 0 and 8.3
    
    
   - Single response:
    
    @99% confidence - we estimate that to observe an individual t-cell repsonse of 0.3, the mass should be between 0 and 21.92g
    

## Residual Scatter:

```{r}
res = resid(object = linear_model)
ggplot() + geom_point(aes(y=res, x=bird_data$Mass)) + xlab('Mass') + ylab('T-Cell Residual') + ggtitle('Residual Plot')

student = rstudent(linear_model)
x = seq(min(student), max(student), length=50)
normal = (dnorm(x))

ggplot() + geom_histogram(aes(x=student, y =..density..), binwidth = 1, frequency=FALSE, fill='LightGreen') + geom_line(aes(x=x, y=normal), color='Red') + ggtitle('Studentized Residuals')

```

## Verify R2
  
```{r}
cor(bird_data$Mass, bird_data$Tcell)^2
```

33.3% of the overall variance in T-cell count is explained by the ability to carry stones
    
## Question 2

```{r}
sleuth_data = Sleuth3::ex0727
write.csv(sleuth_data, 'sleuth_data.csv')
lm_sleuth = lm(data=sleuth_data, formula = Tcell ~ Mass)
summary(lm_sleuth)
confint(lm_sleuth, level=0.99)

```

## See Attached Excel for Rest of Analysis
  