---
title: "Unit 10 Takeaways"
author: "Sean Kennedy"
date: "2/29/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(orcutt)
library(tidyverse)
```

```{r 10.3.3}

bonds = read.csv('10_year_bond_rate_2010-2015.csv')
p = plotts.sample.wge(bonds$Adj.Close)
```
```{r 10.3.4}

bonds_t = artrans.wge(bonds$Adj.Close, phi.tr = c(1))
p = plotts.sample.wge(bonds_t)

```

```{r 10.3.5}


aic5.wge(bonds_t)

```

```{r 10.3.6}


aic5.wge(bonds$Adj.Close)

```

```{r 10.7.3}

ar6 = est.ar.wge(bonds$Adj.Close, p=6, type='burg')
ar8 = est.ar.wge(bonds$Adj.Close, p=8, type='burg')

```

```{r 10.7.5}
data = read.csv('zero_one_or_tworootsofone.csv')


ar6 = est.ar.wge(data$X, p=6, type='burg')
ar8 = est.ar.wge(data$X, p=8, type='burg')

```

```{r 10.8.2}
library(tseries)
 x = gen.arma.wge(200,phi = c(.9), sn = 5)

 
adf.test(x)

```


```{r 10.8.4}

p_vals = c()
test_stats = c()
for(i in seq(1:100)){

  x = gen.arma.wge(200,phi = c(.9))
  p_val = adf.test(x)$p.value
  t_stat = adf.test(x)$statistic
  p_vals = append(p_vals, p_val)
  test_stats = append(test_stats, t_stat)

}
all_runs = data.frame(p_vals, test_stats)

ftr = all_runs %>% filter(p_vals >= 0.05)

length(ftr$p_vals)/length(all_runs$p_vals)
```

```{r 10.9.4}
airline = read_csv('swadelay.csv')

a12_est = est.ar.wge(airline$arr_delay, p = 15, type = 'burg' )
 
```

```{r 10.10.3}
d1 = artrans.wge(airline$arr_delay, phi.tr = c(1))
d1.12 = artrans.wge(d1, c(rep(0, 11), 1))
#plotts.sample.wge(airline$arr_delay)
ar12 = est.ar.wge(d1.12, p= 12, type = 'burg')

```


```{r 10.12.2}

#Step 2: Fit a regression line to the data using the following code:
p_vals = c()
for(i in seq(0:10)){
t = seq(1,100,1)
x = gen.sigplusnoise.wge(100, b0 = 0, b1= 0, phi= .95)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
p_vals = append(p_vals, summary(fit)$coefficients[,4]) 
}
data.frame(p_vals)

p_vals = c()
for(i in seq(0:10)){
t = seq(1,100,1)
x = gen.sigplusnoise.wge(100, b0 = 0, b1= 0, phi= .99)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
p_vals = append(p_vals, summary(fit)$coefficients[,4]) 
}
data.frame(p_vals)
```





```{r 10.12.5}

x = gen.sigplusnoise.wge(100, b0 = 0, b1= 0, phi= .95)  
#note that there is not a seed this time. We will generate a different realization each time.
t = seq(1,100,1)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
summary(fit) # record whether it rejected or failed to reject. 

```




```{r 10.13.3}
x = seq(1, length(airline$arr_delay), 1)
y = airline$arr_delay
fit = lm(x~y, data=data.frame(x=x, y=y))
summary(fit)
```

```{r 10.13.5}
coc = cochrane.orcutt(fit)
```

### FOR LIVE SESSION

```{r FOR_LIVE}

pzza = read.csv('pzza.csv')
p = plotts.sample.wge(pzza$price)
p.1 = artrans.wge(pzza$price, phi.tr = c(1))
p = plotts.sample.wge(p.1)

```

#### ARIMA OVERFIT

```{r}
#Papa john's stock price
ar12 = est.ar.wge(pzza$price, p=12, type='burg')
aic5.wge(pzza$price, p=0:12, type = 'aic')
plotts.sample.wge(pzza$price)
# No evidence of monthly seasonality perhaps ar1
aic5.wge(pzza$price, p=0:12, type = 'bic')

```

#### SEASONAL

```{r}
ar12_real = gen.aruma.wge(n=100, s=12, sn=23)
ar12_real_est = est.ar.wge(ar12_real, p=12, type='burg')
ar12 = est.ar.wge(pzza$price, p=12, type='burg')

# No evidence of monthly seasonality

```

#### COCHRANE ORCUTT

```{r}
x = seq(1, length(pzza$price), 1)
y = pzza$price
fit = lm(x~y, data=data.frame(x=x, y=y))
summary(fit)
cochrane.orcutt(fit)

```


### TAKEAWAYS

### TAKEAWAY HYPOTHESIS TESTING

- Null hypothesis H0 is that b = 0 (no trend), when we fail to reject the null, we are accepting that b is non-zero (evidence of a trend). When p-value is less than alpha we **reject the null** which is bad.... because the data we are generating is generated from models where b is very close to zero. Type 2 error - predicting evidence of a trend when there is none (**incorrectly rejecting the null when it's true**)

### TAKEAWAY OLS IS NOT SUFFICIENT :

- OLS Method for S/N is not good, fails to reject at too high of a rate. Terrible way to test - assumptions of OLS are not met by correlated time series data.


### TAKEAWAY COCHRANE ORCUTT (S/N Test):

- Searches for a linear signal (mean reversion)

- fit a linear regression to get estimates of a and b paramters. 

- transform by caculating the difference between the LM and the realization to estimate the correlated residuals

- fit AR1 to get phi_1

### TAKEAWAY SOMETHNG WRONG WITH C-O TEST :

- Fails to converge....

### TAKEAWAY DIFFERENCING TO TEST FOR UNIT ROOTS:

- By diffing data at k lags we can test for the presence of roots Bk 

- Afterwards, we can rinse repeat until we get white noise (hopefully)


