---
title: "Sean_Kennedy_HW6_DDS"
author: "Sean Kennedy"
date: "6/17/2019"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(sqldf)
library(ggplot2)
load('data.RData')
```

## Question 1
 # Part a/b)
```{r}
states = unique(trimws(mh2015_puf$LST))
```
 # Part c)
```{r}
mh2015_puf$LST = trimws(mh2015_puf$LST)
non_mainland = c('AK', 'HI', 'VI', 'GU', 'PR', 'AS')
filtered = mh2015_puf[!(mh2015_puf$LST %in% non_mainland), ]
mainland_states = unique(filtered$LST)
length(mainland_states) # dc included
```
  # Part d)

```{r}
filtered_counts = sqldf("SELECT LST, COUNT(*) as Total FROM filtered GROUP BY LST")
state_plot = ggplot(filtered_counts, aes(x=filtered_counts$LST, y=filtered_counts$Total, fill=filtered_counts$LST)) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'State', y='# Facilities', fill='State') 
print(state_plot + ggtitle('Facilities by State') +
  theme(plot.title = element_text(hjust = 0.5)))
```
  
## Question 2
```{r}
state_data = read.csv('~/Documents/SMU/DoingDataScience/Homework/HW6/statesize.csv')
merged_data = sqldf("SELECT 
                      StateName
                      , Region
                      , SqMiles
                      , SUM(Total) as FacilityCount
                    FROM filtered_counts f 
                    LEFT JOIN state_data s ON (CASE WHEN f.LST = 'DC' THEN 'VA' ELSE f.LST END) = s.Abbrev
                    GROUP BY StateName, Region, SqMiles
                    ")
merged_data$FacilitiesPer1KSqMile = merged_data$FacilityCount/merged_data$SqMiles/1000

state_plot_by_sq_mile = ggplot(merged_data, aes(x=reorder(merged_data$StateName,merged_data$FacilitiesPer1KSqMile) , y=merged_data$FacilitiesPer1KSqMile, fill=merged_data$Region)) + geom_bar(stat = "identity") + coord_flip() + labs(x = 'StateName', y='# Facilities/Sq Mile', fill='Region') 
print(state_plot_by_sq_mile + ggtitle('Facilities/sq mile by State') +
  theme(plot.title = element_text(hjust = 0.5)))
```
 
 "Facilities seem to be highly concentrated in the Northeast and very few facilities in the West (save for California).
  Further analysis should be undertaken to study the number of veterans that are in each state/region. Simply assuming that the
 geographic size of state is a good measure of the requisite number of VA hospitals is inadequate. Larger states could have larger
 populations, but that is not a gurantee. This analysis would be better done by analyzing the number of hospitals per 1000 veterans, not 
 per 1000 sq miles."
