---
title: "Unit 12 Takeaways"
author: "Sean Kennedy"
date: "3/24/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(orcutt)
library(tidyverse)
library(vars)
```

#```{r}
#TODO: write this function
 lag_all <-- function(tibble){
   
    tibble %>% dplyr::summarise_all(class) %>% tidyr::gather(variable, class)
   
 }
#```

```{r 12.2.5 12.5.2-5}

bus = read.csv('businesssales.csv')

#bus_lag = lag(bus, 1)
X = cbind(bus$sales, bus$ad_tv, bus$ad_online, bus$discount)
VARselect(X, lag.max=10, type='const', season=NULL, exogen=NULL)

fit = VAR(X, p=2, type='const')

VAR(X, p=2, type='trend')

ccf(bus$ad_tv, bus$sales)

AIC(fit)
```

```{r 12.3.4}
lag_data = read.csv('whatisthelag.csv')

ccf(lag_data$Y, lag_data$X1)

```

```{r FOR_LIVE_SK}

CM = read.csv('la_cmort_study.csv')

p = plotts.sample.wge(CM$cmort)

#appears seasonal, yearly from spectral d
CM_52 = artrans.wge(CM$cmort, phi.tr = c(rep(0, 51), 1))

#model the residuals
aic5.wge(CM_52, type='bic') #selects 2,0
p_val_24 = ljung.wge(CM_52)$pval
p_val_48 = ljung.wge(CM_52, K = 48)$pval #barely rejects
acf(CM_52,lag.max = 48)
c(p_val_24, p_val_48) # Fail to reject null (null is that rho_k are all uncorrelated)

#seasonal model predictions
predstemp = fore.aruma.wge(CM$temp,s = 52, n.ahead = 20)
predspart = fore.aruma.wge(CM$part,s = 52, n.ahead = 20)


# Model cmort based on predicted part and temp using MLR with Cor Erros (cor errors b/c time is included?)
ksfit = lm(cmort~temp+part+Week, data = CM)
phi = aic.wge(ksfit$residuals)

fit = arima(CM$cmort,order = c(phi$p,0,0), seasonal = list(order = c(1,0,0), period = 52), xreg = cbind(CM$temp, CM$part, CM$Week))

# Check for whiteness of residuals
acf(fit$residuals) #appear white, but still FTR for the test
ljung.wge(fit$residuals) # pval = .048
ljung.wge(fit$residuals, K = 48) # pval = .002


#load the forecasted Part and Temp in a data frame
next20 = data.frame(temp = predstemp$f, part = predspart$f, Week = seq(509,528,1))
#get predictions
predsCMort = predict(fit,newxreg = next20)
#plot next 20 cmort wrt time
plot(seq(1,508,1), CM$cmort, type = "l",xlim = c(0,528), ylab = "Cardiac Mortality", main = "20 Week Cardiac Mortality Forecast")
lines(seq(509,528,1), predsCMort$pred, type = "l", col = "red")

#Find ASE  Need to forecast last 30 of known series.  
CMsmall = CM[1:478,]
ksfit = lm(cmort~temp+part+Week, data = CMsmall)
phi = aic.wge(ksfit$residuals)
attach(CMsmall)

#getting an error here...
#Error in optim(init[mask], armafn, method = optim.method, hessian = TRUE, : initial value in 'vmmin' is not finite
fit = arima(cmort,order = c(phi$p,0,0), seasonal = list(order = c(1,0,0), period = 52), xreg = cbind(temp, part, Week))

last30 = data.frame(cmort = CM$cmort[479:508], part = CM$part[479:508], Week = seq(479,508,1))
#get predictions
predsCMort = predict(fit,newxreg = last30)


ASE = mean((CM$cmort[479:508] - predsCMort$pred)^2)
ASE

#X = cbind(CM$temp, CM$part, CM$cmort)
#VARselect(X, lag.max=20, type='const', season=NULL, exogen=NULL)

#fit_const = VAR(X, p=2, type='const')
#fit_const
#VAR(X, p=2, type='trend')

#ccf(bus$ad_tv, bus$sales)

#AIC(fit)
```

### TAKEAWAYS


### MULTIVARIATE REGRESSION

n = number of observations
m = number of features (independent variables)

### ARIMA FUNCTION

- Will estimate the fit of the independent variables
- Returns the constant
- Returns the AR parameters

### DOC - YOUR HAIR IN 12.2.3 LOOKS..... BAD

- Clean it up :)

### USE LAG VARIABLES WHEN THEIR OCCURENCE IN THE PAST INFLUENCES THE VALUES YOU ARE TRYING TO PREDICT

### CROSS CORRELATION

- Same as autocorrelation, but amongst different variables $X_i, X_j...X_m$
  
- Can be used to search for leading/lagging indicators

- ccf() function lags the y variable (second position)


### VECTOR AR(P)

$X_{t1}$ refers to the observation of variable $X_1$ at time t

p=2

$X_{t1}  = \beta_1 + \phi_{11(1)}X_{t-1, 1} + \phi_{12(1)}X_{t-1, 2} + \phi_{11(2)}X_{t-2, 1} + \phi_{12(2)}X_{t-2, 2} + a(t)_1$


$X_{t2}  = \beta_2 + \phi_{11(1)}X_{t-1, 1} + \phi_{12(1)}X_{t-1, 2} + \phi_{11(2)}X_{t-2, 1} + \phi_{12(2)}X_{t-2, 2} + a(t)_1$


### FORECASTS


### VARSelect

- Used to determine the proper lag

- used in tandem with VAR

