---
title: "Time Series Unit 4"
author: "Sean Kennedy"
date: "1/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
library(tidyverse)
library(stats)
library(lubridate)
```

```{r EXAMPLE_QUESTION}
factor.wge(phi = c(1.59, -.544, -.511, .222))
```

```{r IN_THE_REAL_WORLD}
## characteristic equation
## 1+.5Z+.6Z^2 = 0
## Root is 1.2909 and is outside the unit circle - hence stationary
## See excel for calculations

factor.wge(phi = c(-0.5, -0.6))

sample = plotts.true.wge(phi = c(-0.5, -0.6))
```

```{r WALMART}
walmart_data = read.csv('Walmart.csv')
item8 = walmart_data %>% filter(item == 50 & store==9)
sample = plotts.sample.wge(item8$sales)
```
### Analysis:

  - Let's check the proposed models for their frequency components
  
```{r ANALYSIS}
phi1 = c(.967)
factor.wge(phi = phi1)
sample = plotts.true.wge(phi = phi1)
## simple wandering model - clearly not sufficient to describe cyclical patterns

phi2 = c(1.452, -.453, -.294, .175, .237, -.154)
factor.wge(phi = phi2)
sample2 = plotts.true.wge(phi = phi2)
# has some wandering but is missing frequencies that are clearly present in sample spectral decomp

phi3 = c(1.445, -.411, -.038, .170, .362, -.245, -.177, .213)
factor.wge(phi = phi3)
#sample3 = plotts.true.wge(phi = phi3)
# AR1 component is non-stationary

phi4 = c(1.384, -.359, -.309, .063, .317, -.140, -.0587, -.199, .2877)
factor.wge(phi = phi4)
sample4 = plotts.true.wge(phi = phi4)
# This one seems to match best to the frequencies observed in the sample, exhibits wandering and has peaks that are close to the observed peaks

```



### TAKEAWAY 1 BACKSHIFT OPERATOR

- Is basically the lag() function in R
- (Bk)Xn = X(n-k) lags X by k periods


### TAKEAWAY 2 CHARACTERISTIC EQ

- Solves determinate of backshift operator matrix
- Simple for AR(1) (solve linear eq)
- Simple for AR(2) (factor or use binomial)
- For AR(p) use R (factor.wge)

### TAKEAWAY 3 STATIONARITY

- ALL roots of characteristic equation must be outside unit circle for model to be stationary
- Factor with largest value of abs recip root will dominate periodic behavior

### TAKEAWAY 4 COMPLEX ROOTS

- Complex roots generate sinusoidial behavior with frequencies between 0 and 0.5
- Required if spectral density has peaks in any other place than 0 or 0.5