---
title: "Sean Kennedy Doing Data Science: HW 3"
output: html_notebook
---

#Question 1: Git Code

(base) Seans-MacBook-Pro:~ skennedy$ cd Documents/

(base) Seans-MacBook-Pro:Documents skennedy$ cd SMU/

(base) Seans-MacBook-Pro:SMU skennedy$ cd DoingDataScience/Homework/HW3/

(base) Seans-MacBook-Pro:HW3 skennedy$ mkdir GIT

(base) Seans-MacBook-Pro:HW3 skennedy$ cd GIT/

(base) Seans-MacBook-Pro:GIT skennedy$ git clone https://github.com/caesar0301/awesome-public-datasets

xcrun: error: invalid active developer path (/Library/Developer/CommandLineTools), missing xcrun at: /Library/Developer/CommandLineTools/usr/bin/xcrun

(base) Seans-MacBook-Pro:GIT skennedy$ xcode-select --install

xcode-select: note: install requested for command line developer tools

(base) Seans-MacBook-Pro:GIT skennedy$ git clone https://github.com/caesar0301/awesome-public-datasets

Cloning into 'awesome-public-datasets'...

remote: Enumerating objects: 24, done.

remote: Counting objects: 100% (24/24), done.

remote: Compressing objects: 100% (10/10), done.

remote: Total 1761 (delta 14), reused 24 (delta 14), pack-reused 1737

Receiving objects: 100% (1761/1761), 665.23 KiB | 1.12 MiB/s, done.

Resolving deltas: 100% (1049/1049), done.

(base) Seans-MacBook-Pro:GIT skennedy$">


```{r echo=TRUE}
titanic_data <- read.table(file = "~/Documents/SMU/DoingDataScience/Homework/HW3/GIT/awesome-public-datasets/Datasets/titanic.csv",header = T,sep = ","); # 
summary(titanic_data)
#females <- titanic_data[which(titanic_data$Survived ==1)]
females <- titanic_data[which(titanic_data$Sex =='female'), ]
hist(females$Age, xlab = 'Age', ylab= 'Count', main = 'Female Passenger Age')
hist(females$Survived, xlab = 'Survived', ylab= 'Count', main = 'Female Passenger Survival')
males <- titanic_data[which(titanic_data$Sex =='male'), ]
hist(males$Age, xlab = 'Age', ylab= 'Count', main = 'Male Passenger Age')
hist(males$Survived, xlab = 'Survived', ylab= 'Count', main = 'Male Passenger Survival')
means = apply(data.matrix(titanic_data[,c('Age', 'Fare', 'Survived')]), MARGIN = 2, FUN=mean, na.rm=TRUE)
real = means == Re(means)
```
```{r}
sleep_data <- read.table(file = "~/Documents/SMU/DoingDataScience/Homework/HW3/sleep_data_01.csv",header = T,sep = ",");
summary(sleep_data)
median_age = apply(data.matrix(sleep_data[,c('Age')]), MARGIN = 2, FUN=median, na.rm=TRUE)
min_sleep = apply(data.matrix(sleep_data[,c('Duration')]), MARGIN = 2, FUN=min, na.rm=TRUE)
max_sleep = apply(data.matrix(sleep_data[,c('Duration')]), MARGIN = 2, FUN=max, na.rm=TRUE)
RSES = c(apply(data.matrix(sleep_data[,c('RSES')]), MARGIN = 2, FUN=mean, na.rm=TRUE), apply(data.matrix(sleep_data[,c('RSES')]), MARGIN = 2, FUN=sd, na.rm=TRUE))
summary = t(summarize(sleep_data))

```
```{r}
summarize =function(x)
{
  report = data.frame(c('Median Age', 'Self Esteem','SE_SD', 'Duration Range'),c(apply(data.matrix(x[,c('Age')]), MARGIN = 2, FUN=median, na.rm=TRUE),apply(data.matrix(x[,c('RSES')]), MARGIN = 2, FUN=mean, na.rm=TRUE)/5, apply(data.matrix(x[,c('RSES')]), MARGIN = 2, FUN=sd, na.rm=TRUE), apply(data.matrix(x[,c('Duration')]), MARGIN = 2, FUN=max, na.rm=TRUE)-apply(data.matrix(x[,c('Duration')]), MARGIN = 2, FUN=min, na.rm=TRUE)))
}
```
```{r}
sets = c(data(package = 'fivethirtyeight'))
sets$results[22,]
df = fivethirtyeight::college_recent_grads
names(df)
length(names(df))
major_count = data.frame(table(df$major_category))
par(las=2)
barplot(major_count$Freq, names.arg=major_count$Var1, horiz=TRUE, main = 'Major Selection',ylab = 'Major', xlab = 'Frequency', col = c('red'))
write.csv(df[names(df)[2:length(names(df))]], file='fivethirtyeight.csv')
```

```{r}
datasets <- read.csv("https://goo.gl/OT8iHa") 
d = datasets[which(datasets$DATAFRAME_NAME =='college_recent_grads'), ]
dim(datasets)
url = d['ARTICLE_URL']
url
d
```







