---
title: "Sean Kennedy HW 12"
author: "Sean Kennedy"
date: "8/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sqldf)
library(tidyquant)
```

```{r}
b_data = read.csv('Brain.csv')
pairs(~ Brain + Body + Gestation + Litter, data=b_data)

```

## Crappy - let's log it!

```{r}
b_data$Brain_log = log(b_data$Brain)
b_data$Body_log = log(b_data$Body)
b_data$Gestation_log = log(b_data$Gestation)
b_data$Litter_log = log(b_data$Litter)

pairs(~ Brain_log + Body_log + Gestation_log + Litter_log, data=b_data)
log_model = lm(data=b_data, formula = Brain_log ~ Body_log + Gestation_log + Litter_log)
summary(log_model)
plot(log_model)
```

## Assumptions of normality, heteroscedascity, linearity and indepedence (?) are all met for the log transformed data

 ln(brain)=0.85+0.58⋅ln(body)+0.42⋅ln(gestation)-0.31⋅ln(litter)
 
Res Plot- slight funnel shape, but not terrible

Stud Res Plot- observations 53 and 25 look suspect

Q-Q: Normal

Histogram of Residuals: also normal


Interpretation


Since this was a log/log model we can only say that a doubling in the input results in a 1.34x multiplicative change in the median brain size. 
