---
title: "Unit 9 Takeaways"
author: "Sean Kennedy"
date: "2/29/2020"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tswge)
```

```{r 9.3.3}

phi= c(0.3, -0.7)
theta = c(-.4)
mu = 37
sigma_a = 4

realization = gen.arma.wge(200, phi=phi, theta=theta, vara=sigma_a, sn=27)
realization = realization + mu
est.arma.wge(realization, p=2, q=1)
mean(realization)
```

```{r 9.5.7}

phi= c(0.3, -0.7)
mu = 37
sigma_a = 4

realization = gen.arma.wge(200, phi=phi, vara=sigma_a, sn=27)
realization = realization + mu
est.ar.wge(realization, p=2, type='burg')

```

```{r 9.8.3}
wn = read.csv('maybewhitenoise1.csv')
plotts.sample.wge(wn$x)

```

```{r 9.8.4}
wn = read.csv('maybewhitenoise2.csv')
plotts.sample.wge(wn$x)

```

```{r 9.9.4}
#CANNOT BE ANSWERED - NO N
```

```{r 9.9.5}
inflation = read.csv('inflation.csv')
plotts.sample.wge(inflation$Inflation)
aic5.wge(inflation$Inflation, type='aic')
aic5.wge(inflation$Inflation, type='bic')
```


```{r 9.12.2}
data = read.csv('armawhatpq1.csv')

acf(data$x)
pacf(data$x)
```

```{r 9.12.4}
data = read.csv('armawhatpq1.csv')
aic5.wge(data$x, type='aic')
aic5.wge(data$x, type='bic')
```

```{r 9.12.3}
data = read.csv('inflation.csv')
acf(data$Inflation)
pacf(data$Inflation)
```

```{r FOR_LIVE_SESSION}
texas = read.csv('texasgasprice.csv')
aic5.wge(texas$Price)

ar2_est_mle = est.ar.wge(texas$Price, p=2, type = 'mle')
ar2_est_burg = est.ar.wge(texas$Price, p=2, type = 'burg')

phi_burg = ar2_est_burg$phi
phi_mle = ar2_est_mle$phi

predict_burg = fore.arma.wge(texas$Price, phi = phi_burg, n.ahead = 24, lastn = FALSE)
predict_mle= fore.arma.wge(texas$Price, phi = phi_mle, n.ahead = 24, lastn = FALSE)

plotts.sample.wge(texas$Price)

plotts.true.wge(500, phi = phi_burg, vara=ar2_est_burg$avar)
plotts.true.wge(500, phi = phi_mle, vara=ar2_est_mle$avar)
## results from burg and MLE are identical?
```

```{r}
phi_burg

```

```{r}
phi_mle

```

### Takewawy 1

  - ALWAYS CHECK FOR WHITE NOISE
  
    - Evidence of constant/mean variance? 
    
### Takeaway 2

  - Yule Walker estitmates are not very useful when roots are close to unit circle (i.e realization exhibits wandering)
  
  - Burg estimates will not produce roots inside unit circle (which is good!)
  
  - MLE estimates are better than YW but can produce roots inside unit cirlcle
  
### Takeaway 3

  - AIC is a more useful metric than RMSE because it penalizes for model complexity
  
## Takeaway 4

  - Interpreting ACF/PACF plots for MA/AR models:
  
    - MA models
        
        -ACF: drop off after k=q
        
        -PACF 
        
    -AR models
    
      - ACF: depends on nature of roots
      
      
      - PACF