---
title: "Sean_Kennedy_Homework_5_Stats"
author: "Sean Kennedy"
date: "June 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggpubr)
library(multcomp)
library(car)
library(sqldf)
```

## R Question 1 Problem 5.25 from Text (ANOVA)

```{r}
raw_data = read.csv('GWarming_gender.csv')
raw_data[,'Gender_1'] = factor(raw_data[,'Gender_1'])
res.aov = aov(Rating_1 ~ Gender_1, data=raw_data)
summary(res.aov)
```

## Let's plot the fuck out of this!

```{r education_data, echo=FALSE}
# tak a random sample for good measure
set.seed(1234)
dplyr::sample_n(raw_data, 10)
levels(raw_data$gleason)
#show some group stats
group_by(raw_data, Gender_1) %>% summarise(
  count = n(),
  mean = mean(Rating_1, na.rm = TRUE),
  sd = sd(Rating_1, na.rm = TRUE)
)
ggboxplot(raw_data, x='Gender_1', y='Rating_1', color = 'Gender_1',
          palette=palette(rainbow(length(levels(raw_data$Gender_1)))),
          ylab = 'Rating', xlab = 'Gender'
          )

ggline(raw_data, x = 'gleason', y = 'psa', 
       add = c('mean_se'), #(other values include: mean_sd, mean_ci, median_iqr,jitter ....)
       ylab='PSA', xlab='Gleason')

ggplot(raw_data, aes(x=psa)) + geom_histogram(binwidth=.5, colour="black", fill="white") + 
    facet_grid(gleason ~ .) 
```

## Tukey Multiple Pairwise Comparisons

```{r}
TukeyHSD(res.aov)
```

## Let's calc the same thing using multcomp package 
  
    #(why doensn't this work?)
```{r}
summary(glht(res.aov, linfct = mcp(gleason = 'Tukey')))

```

## pariwise t-test, because "Why not?"
  
  # creates a matrix of t-tests between classes
```{r}
pairwise.t.test(raw_data$psa, raw_data$gleason,
                p.adjust.method = 'BH')

```

## Check homogeneity of variance assumption
```{r}
plot(res.aov, 1)
```

## Levene's test for homogeneity of variance
```{r}
leveneTest(psa ~ gleason, data=raw_data)
```

## Relaxed Homogeneity Tests
  
  - Welch one-way
  
  - paired t-test w/o equality of sd assumption)

```{r}
oneway.test(psa ~ gleason, data = raw_data)
pairwise.t.test(raw_data$psa, raw_data$gleason, p.adjust.method = 'BH', pool.sd = FALSE)
```

## Check Normality Assumption

```{r}
# Q-Q plot of residuals
plot(res.aov, 2)

# Extract the residuals 
aov_residuals = residuals(object = res.aov)

#Shapiro-Wilke on residuals
shapiro.test(x = aov_residuals)
```

## Non Parametric Tests (Kruskal-Wallis Ranksum Test)

```{r}
kruskal.test(psa ~ gleason, data=raw_data)

```

